### 2013-03-01 branching-transit.tex
WorkingWiki importation

Suppose we have branching processes beginning at the origin at some constant rate $\lambda$ in time.
Then the occupation density at distance $x$, i.e.\ the expected number of particles found at location $x$ at any given time, at stationarity,
is $\lambda \int_0^\infty \E[Z_t(dx)] dt$, where $Z_t$ is the branching process.
Since spatial motion is independent of branching,
\begin{align}
  \E[Z_t(dx)] = e^{-\alpha t} \frac{1}{(2\pi \sigma^2 t)^{d/2}} \exp\left(-\frac{x^2}{2\sigma^2 t} \right) ,
\end{align}
and so using G\&R 3.471.9 and 8.468 again,
\begin{align} \label{eqn:bp_occupation}
  \int_0^\infty \E[Z_t(dx)] dt  &= \int_0^\infty e^{-\alpha t} \frac{1}{(2\pi \sigma^2 t)^{d/2}} \exp\left(-\frac{x^2}{2\sigma^2 t} \right) dt \\
  &= \begin{cases}
    \frac{ 2^{1/4} x^{1/2} }{ \sqrt{\pi} \sigma^{3/2} \alpha^{1/4} } K_{1/2} \left(\frac{x \sqrt{2\alpha}}{\sigma}\right) =
    \frac{1}{\sigma \sqrt{2 \alpha}} \exp\left( - \frac{ \sqrt{2\alpha} x }{ \sigma } \right) \qquad & d=1 \\
    \frac{1}{\sigma^2 \pi} K_0\left(\frac{x \sqrt{2\alpha}}{\sigma}\right) \sim \frac{ e^{-\frac{x\sqrt{2\alpha}{\sigma}}} }{ \sqrt{ \sigma^3  2\pi x \qquad & d=2  ,
\end{cases} 
\end{align}
where $K_0(z)$ is the modified Bessel function of the second kind of order 0.
Note that this is the Green function for a Brownian motion killed at rate $\alpha$.

A more back-of-the-envelope calculation of this would be:
we know that the probability a branching process arrives at the patch is given by $(1/\mu_K)\exp(-x\sqrt{2 \alpha}/\sigma)$;
if it does it is present in $\mu_K$ copies, each of which leave a total of $1/s_m$ offspring.
To translate this to the above we need the typical area covered by these (in some sense?), 
which must be around $\sigma \sqrt{2 \alpha}$.

@@

For reference, by G\&R 3.471.9 and 8.468,
\begin{align}
  \int_0^\infty x^{n+1/2} e^{-\frac{\beta}{x}-\gamma x} dx &= 2 \left( \frac{\beta}{\gamma} \right)^{\frac{2n+1}{4}} K_{n+3/2}(2\sqrt{\beta\gamma}) \quad \mbox{(for general $n$)} \\
  \mbox{($n$ an integer)} \quad &= \sqrt{\pi} \beta^{\frac{n}{2}} \gamma^{-\frac{n+1}{2}} e^{-2\sqrt{\beta\gamma}} \sum_{k=1}^n \frac{ (n+k)! }{ k! (n-k)! (4\sqrt{\beta \gamma})^k } .
\end{align}

Note that (8.451.6) $K_\nu(z) = \sqrt{\frac{\pi}{2z}} e^{-z} (1 + O(1/z))$.

\section{The big picture}

@@

We can calculate the rate of ``core'' establishment, i.e.\ the mean number of reproductive events that lead to establishment as defined below, in two ways:
\begin{align}
  (\mbox{rate of establishment}) &= (\mbox{outflux of BPs from old patch}) \times (\mbox{\# estab events per BP}) \\
    \begin{split} &= (\mbox{outflux of BPs from old patch}) \times (\mbox{prob of estab of a distant BP}) \\
  & \qquad \qquad \times (\mbox{\# estab events per successful BP}) \end{split} \\
  \begin{split} &= (\mbox{subcritical occupation density}) \times \rho \times (\mbox{\# offspring per subcritical indiv}) \\
    & \qquad \qquad \times (\mbox{prob of estab in the patch}) . \end{split} 
\end{align}
The probability of establishment of a branching process is in equation~\eqref{eqn:bp_prob_estab};
the subcritical occupation density is the integral of equation~\eqref{eqn:eqfreq} over the patch;
the number of (extra) offspring per subcritical individual is $s_B+s_m$;
and the probability of establishment in the patch is $2s_B/\xi^2$.
The ``outflux of branching processes'' we have not computed;
but we have compute the occupation density from this branching process approximation above,
so equating ($\rho$ times $C\exp(-\sqrt{2 s_m}x/\sigma)$) and (outflux times equation~\eqref{eqn:bp_occupation})
we have that the effective outflux (from the edge of the already occupied patch!) is 
\begin{align}
  (\mbox{outflux of BPs from old patch}) = \frac{ C \rho \sqrt{\alpha} }{ \sqrt{\pi} } ,
\end{align}
rather than $C\rho$ which we might have guessed.
The number of establishment events per BP is as follows, in averages: 
with prob $1/\mu_K \exp(-\sqrt{\alpha} x/ \sigma)$ a total of $\mu_K$ indivs arrive at the new patch; 
each produce $1/s_m$ total core offspring;
each of these produce $s_B+s_m$ extra offspring which each have chance $2s_B/\xi^2$ of establishing,
i.e.
\begin{align}
  (\mbox{\# estab events per BP}) = \frac{(s_B+s_m) 2 s_B }{ s_m \xi^2 } \exp\left(-\frac{\sqrt{\alpha} x}{\sigma}\right) .
\end{align}
\textbf{Now} note that what we \textbf{actually} want to compute is not this rate, but rather 
\begin{align}
  \migrate &= (\mbox{outflux of BPs from old patch}) \times (\mbox{prob of estab of a distant BP}) \\
  &= C \rho \sqrt{\alpha} \times \frac{2 s_b}{\sqrt{\pi} \xi^2} \exp\left(- \frac{\sqrt{\alpha}x}{\sigma} \right) .
\end{align}
Finally, combining these, we can also compute that
\begin{align}
  (\mbox{\# estab events per successful BP}) = \frac{s_B+s_m}{ \sqrt{s_m} } .
\end{align}

>>>>

Suppose we have branching processes beginning at the origin at some constant rate $\lambda$ in time.
Then the occupation density at distance $x$, i.e.\ the expected number of particles found at location $x$ at any given time, at stationarity,
is $\lambda \int_0^\infty \E[Z_t(dx)] dt$, where $Z_t$ is the branching process.
Since spatial motion is independent of branching,
\begin{align}
  \E[Z_t(dx)] = e^{-\alpha t} \frac{1}{(2\pi \sigma^2 t)^{d/2}} \exp\left(-\frac{x^2}{2\sigma^2 t} \right) ,
\end{align}
and so using G\&R 3.471.9 and 8.468 again,
\begin{align} \label{eqn:bp_occupation}
  \int_0^\infty \E[Z_t(dx)] dt  &= \int_0^\infty e^{-\alpha t} \frac{1}{(2\pi \sigma^2 t)^{d/2}} \exp\left(-\frac{x^2}{2\sigma^2 t} \right) dt \\
  &= \begin{cases}
    \sqrt{ \frac{ 2 x }{ \pi \sigma^{3} \sqrt{2\alpha} } }  K_{1/2} \left(\frac{x \sqrt{2\alpha}}{\sigma}\right) =
    \frac{1}{\sigma \sqrt{2 \alpha}} \exp\left( - \frac{ x \sqrt{2\alpha} }{ \sigma } \right) \qquad & d=1 \\
    \frac{1}{\sigma^2 \pi} K_0\left(\frac{x \sqrt{2\alpha}}{\sigma}\right) 
    \sim \frac{ 1 }{ \sqrt{ \sigma^3  2\pi x \sqrt{2\alpha} } } \exp\left(-\frac{x\sqrt{2\alpha}}{\sigma}\right) \qquad & d=2  ,
\end{cases} 
\end{align}
where $K_0(z)$ is the modified Bessel function of the second kind of order 0.
Note that this is the Green function for a Brownian motion killed at rate $\alpha$.

A more back-of-the-envelope calculation of this would be:
we know that the probability a branching process arrives at the patch is given by $(1/\mu_K)\exp(-x\sqrt{2 \alpha}/\sigma)$;
if it does it is present in $\mu_K$ copies, each of which leave a total of $1/s_m$ offspring.
To translate this to the above we would need the typical area covered by these (in some sense?).
% which must be around $\sigma \sqrt{2 \alpha}$.

@@

For reference, by G\&R 3.471.9 and 8.468,
\begin{align}
  \int_0^\infty x^{n-1/2} e^{-\frac{\beta}{x}-\gamma x} dx &= 2 \left( \frac{\beta}{\gamma} \right)^{\frac{2n+1}{4}} K_{n+1/2}(2\sqrt{\beta\gamma}) \quad \mbox{(for general $n$)} \\
  \mbox{($n$ an integer)} \quad &= \sqrt{\pi} \beta^{\frac{n}{2}} \gamma^{-\frac{n+1}{2}} e^{-2\sqrt{\beta\gamma}} \sum_{k=1}^n \frac{ (n+k)! }{ k! (n-k)! (4\sqrt{\beta \gamma})^k } .
\end{align}

Note that (8.451.6) $K_\nu(z) = \sqrt{\frac{\pi}{2z}} e^{-z} (1 + O(1/z))$ for all $\nu$, and $K_{-\nu}(z)=K_\nu(z)$.

We have $n-1/2 = - d/2$, $\beta=x^2/2\sigma^2$, and $\gamma = \alpha$, leading to
\begin{align}
\frac{1}{(2\pi\sigma^2 )^{d/2}} \int_0^\infty x^{n-1/2} e^{-\frac{\beta}{x}-\gamma x} dx \sim
\frac{ 1 }{ (2\pi)^\frac{d-1}{2} \sigma^d } \left( \frac{\sigma}{x} \right)^{\frac{d-1}{2}} ({2\alpha})^\frac{3-d}{4} \exp\left( - \frac{ x \sqrt{2 \alpha} }{ \sigma } \right) \quad \mbox{as}\; \frac{x\sqrt{2\alpha}}{\sigma} \to \infty .
\end{align}

%%%%%%%%%
\section{The big picture}

@@

We can calculate the rate of ``core'' establishment, i.e.\ the mean number of reproductive events that lead to establishment as defined below, in two ways:
\begin{align}
  (\mbox{rate of establishment}) &= (\mbox{outflux of BPs from old patch}) \times (\mbox{\# estab events per BP}) \\
    \begin{split} &= (\mbox{outflux of BPs from old patch}) \times (\mbox{prob of estab of a distant BP}) \\
  & \qquad \qquad \times (\mbox{\# estab events per successful BP}) \end{split} \\
  \begin{split} &= (\mbox{subcritical occupation density}) \times \rho \times (\mbox{\# offspring per subcritical indiv}) \\
    & \qquad \qquad \times (\mbox{prob of estab in the patch}) . \end{split} 
\end{align}
The probability of establishment of a branching process is in equation~\eqref{eqn:bp_prob_estab};
the subcritical occupation density is the integral of equation~\eqref{eqn:eqfreq} over the patch;
the number of (extra) offspring per subcritical individual is $s_B+s_m$;
and the probability of establishment in the patch is $2s_B/\xi^2$.
The ``outflux of branching processes'' we have not computed;
but we have computed the occupation density from this branching process approximation above,
so equating ($\rho$ times $C\exp(-\sqrt{2 s_m}x/\sigma)$) and (outflux times equation~\eqref{eqn:bp_occupation})
we have that the effective outflux (from the edge of the already occupied patch!) is 
\begin{align}
  (\mbox{outflux of BPs from old patch}) = C \rho \sigma \sqrt{2 \alpha} ,
\end{align}
rather than $C\rho$ which we might have guessed.
The number of establishment events per BP is as follows, in averages: 
with prob $1/\mu_K \exp(-\sqrt{\alpha} x/ \sigma)$ a total of $\mu_K$ indivs arrive at the new patch; 
each produce $1/s_m$ total core offspring;
each of these produce $s_B+s_m$ extra offspring which each have chance $2s_B/\xi^2$ of establishing,
i.e.
\begin{align}
  (\mbox{\# estab events per BP}) = \frac{(s_B+s_m) 2 s_B }{ s_m \xi^2 } \exp\left(-\frac{\sqrt{\alpha} x}{\sigma}\right) .
\end{align}
\textbf{Now} note that what we \textbf{actually} want to compute is not this rate, but rather 
\begin{align}
  \migrate &= (\mbox{outflux of BPs from old patch}) \times (\mbox{prob of estab of a distant BP}) \\
  &= C \rho \sqrt{\alpha} \times \frac{2 s_b}{\sqrt{\pi} \xi^2} \exp\left(- \frac{\sqrt{\alpha}x}{\sigma} \right) .
\end{align}
Finally, combining these, we can also compute that
\begin{align}
  (\mbox{\# estab events per successful BP}) = \frac{s_B+s_m}{ \sqrt{s_m} } .
\end{align}
